---
title: "ch3"
author: "Chirag Lakhani"
date: "April 15, 2015"
output: html_document
---

This dataset is the Heart and Estrogen/Progestin Study (HERS), a clinical trial of hormone therapy for prevention of recurrent heart attacks and deaths among 2,763 post-menopausal women with existing coronary heart disease. 

We will first calculate a t-test to compare the difference in glucose levels based on excercise
```{r}
library(ggplot2)
hersdata <- read.delim("hersdata.txt")
head(hersdata[,c('glucose','exercise')])
ggplot(hersdata, aes(x=glucose, fill=exercise)) + geom_bar(position="dodge") + ggtitle("Distribution of glucose levels for both exercise groups")

qplot(exercise, glucose, data=hersdata, geom=c("boxplot"), 
   fill=exercise, main="Boxplot of Glucose Level by Exercise",
   xlab="", ylab="Glucose Level")

```

Now we will calculate the t-test for average glucose values by exercise 
```{r}
exerciseYes <- hersdata[hersdata$exercise=='yes',]
exerciseNo <- hersdata[hersdata$exercise=='no',]
t.test(exerciseYes$glucose,exerciseNo$glucose)
```